\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-08-01 manuscript.Rnw
Added section in discussion about choice to use R6

%% - \Abstract{} almost as usual
\Abstract{
Health economic models simulate the costs and effects of health technologies for use in health technology assessment to inform efficient use of scarce resources. Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited. \pkg{hesim} is an \proglang{R} package that helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner. Supported model types include cohort discrete time state transition models, individual continuous time state transition models, and partitioned survival models, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and utility in a flexible way. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs. Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible. Decision analysis within a cost-effectiveness framework is performed using simulated costs and quality-adjusted life years from a PSA.
}

@@

A third feature worth mentioning is that a PSM can be extended to $N$ health states. The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy. One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time spent (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) after progression. An example four-state model is provided on the package website.

@@

\pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling. As noted in the introduction, there are some important differences between \pkg{hesim} and \pkg{heemod} related to the supported model types and integration of statistical and economic modeling. In addition, as with \pkg{mstate}, simulations are not vectorized which could result in significantly reduced performance. To test this, we simulated a time-inhomogeneous total hip replacement example from \citet{briggs2006decision}. A cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.

>>>>

%% - \Abstract{} almost as usual
\Abstract{
Health economic models simulate the costs and effects of health technologies for use in health technology assessment to inform efficient use of scarce resources. Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited. \pkg{hesim} is an \proglang{R} package that helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner. Supported model types include cohort discrete time state transition models, individual continuous time state transition models, and partitioned survival models, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. A modular design allows users to combine separate submodels for disease progression, costs, and utility in a flexible way. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs. Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible. Decision analysis within a cost-effectiveness framework is performed using simulated costs and quality-adjusted life years from a PSA.
}

@@

A third helpful feature that we did not mention is that a PSM can be extended to $N$ health states. The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy. One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time spent (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) after progression. An example four-state model is provided on the package website.

A design decision that is worth commenting on is the use of \pkg{R6}. While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely. It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that we can have separate submodels for disease progression, costs, and utility, which can be combined into an economic model. A particularly important benefit of \pkg{R6} is that methods can belong to objects. In our framework, this means that submodels can not only contain inputs for the economic model, but can also generate outputs. For example, the disease model of a PSM not only contains the parameters of the survival models and the input data required for prediction, but can also generate relevant outcomes such as survival curves or hazard functions. Likewise, the disease model of an iCTSTM not only stores parameters and input data, but can be used to simulate trajectories through a multi-state model and compute state probabilities. This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way. The latter may be especially relevant to users who only want to use some features of the package; for, instance a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels. All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.

@@

\pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling. As noted in the introduction, there are some important differences between \pkg{hesim} and \pkg{heemod} related to the supported model types and integration of statistical and economic modeling. In addition, as with \pkg{mstate}, simulations are not vectorized, which could result in significantly reduced performance. To test this, we simulated a time-inhomogeneous total hip replacement example from \citet{briggs2006decision}. A cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.
